What's hot and what's not:\\Windowed developer topic analysis

### 2009-06-17 lda-paper.tex
more edits and latex8 updates

\documentclass[times, 10pt,twocolumn]{article} 
\usepackage{latex8}

\usepackage{graphicx,amsmath,comment,fullpage}
\newcommand{\igH}[1]{\includegraphics[height=.9\textheight]{#1}}
\newcommand{\igW}[1]{\includegraphics[width=.9\textwidth]{#1}}
\newcommand{\igWh}[1]{\includegraphics[width=.7\textwidth]{#1}}
\newcommand{\igWhalf}[1]{\includegraphics[width=.45\textwidth]{#1}}
\newcommand{\lda}{Latent Dirichlet Allocation}

@@

\maketitle
\begin{abstract}

@@

\section{Introduction}

We care about the topics of development. We want to know given a
certain period of time, what were the popular topics being worked on
during that period. We also want to know if any of those topics
occurred again, whether before or after our current period.

These topics could help stakeholders determine what their coworkers
were working on, what were common issues, what requirements were being
worked on, etc.

Topic analysis extracts groups of words from documents (commit log
comments) that are word distributions, topics, which characterize
clusters of documents. These extracted topics often correlate with
actual development topics that the developers discussed.

Topic analysis often allows for a single document, such as a commit
message, to be related to multiple topics. This maps well to
commits which often have more than one purpose.  A topic represents
both a word distribution and a group of commit log comments that are
related to each other by their content.  In this case a topic is a set
of tokens.

@@

Figure \ref{fig:lda} illustrates the kind of visualization we want to
automatically generate. We want to see topics that are unique to a
certain time period, displayed across the time axis. Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a single word drawn from
itself.

In this paper we explore how topics shift over time in the source
control system (SCS) of a project, using several open source database
systems as examples. We analyze commit log comments in each time
window and we extract the topics of that time window. We expect that
topics will change over time and the similarities and differences in
these topics will indicate developer focus and changes in developer
focus as it changes over time. We will show that most topics are
locally relevant and thus a windowed analysis can allow for both
different kinds of global and local analysis than global topic
analysis would.

Our contributions include:
\begin{itemize}
\item Demonstrating that windowed topic analysis most topics are local
  in scope, thus would be missed by a global topic analysis.
\item Multiple visualizations of topics and trends over time to aid
  communication and analysis of topics and trends.
\item An exploratory case study using these techniques on several open
  source database systems.
\end{itemize}

@@

%%%   \section{Introduction}

@@

\section{Background}

@@

%\subsection{Intro to LDA}

@@

\shrinkit
\subsection{LDA, LSI and Semantic Clustering}
\shrinkit

@@

\shrinkit
\section{Preliminary Case Study}
\shrinkit

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
examined the top 10 most frequent words in that topic.  We found there
were common words across topic clusters, such as \emph{diffs},
\emph{annotate} and \emph{history}; these words probably should be
viewed as stop words. There were notable transitional topics such as
in the first window the word \emph{BitKeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of BitKeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with only a little familiarity with the code base that
naming the topic was straightforward.

@@

\shrinkit
\section{Methodology}
\shrinkit

@@

\shrinkit
\subsection{Extraction of  Repositories and Documents}
\shrinkit

@@

\shrinkit
\subsection{Windowed Sets}
\shrinkit

@@

\shrinkit
\subsection{Apply Topic Analysis to each Window}
\shrinkit

@@

%\subsection{Extract Clusters}

%   6. [ ] Cluster Similarity [0/1]
\shrinkit
\subsection{Topic Similarity}
\shrinkit
%      1. [ ] need the tool
%             grab from the cluster analysis
%   7. [ ] Continuous Cluster Analysis

@@

\shrinkit
\subsection{Visualization}
\shrinkit
%       1. [ ] need the visualizer

@@

\shrinkit
\section{Results}
\shrinkit

@@

%\subsection{MySQL}
%XXX DO MYSQL?
\shrinkit
\subsection{PostgreSQL}
\shrinkit
%   3.   mysql

@@

\shrinkit
\subsection{Firebird}
\shrinkit
%   5.   what's that other db?

@@

\shrinkit
\subsection{MaxDB 7.500}
\shrinkit

@@

The query optimizer was a topic that recurred during MaxDB's
development. In our plots, topics that mention \emph{optimizer} occur
four times, yet in the global-trend-view (Figure
\ref{fig:statictopics}) it is not in any of the topics. A query
optimizer is an important component of an DBMS, but as we have shown
it does not appear as a topic on its own. We tried to remove words to
see if we could get an \emph{optimizer} topic. After removing stop
words and then two of the most common words, the global analysis
finally had a topic with optimizer in its topic ten words. Our analysis shows
that optimizer was important but it had been obscured, but would have
been noticed using the more local topic analysis.

@@

\shrinkit
\subsection{Compare with topics over the entire development}
\shrinkit

@@

%\subsection{Validation}

%   1. well i'm stumped
%   2. look at it?
%   3. investigate those revisions and check the proportion etc
%   4. exploratory
%   5.   work it out
%8.   Validity Threats [0/5]
\shrinkit
\section{Validity Threats}
\shrinkit

@@

\shrinkit
\section{Conclusions}
\shrinkit
%11.   Start file file:/home/abez/projects/lda-paper/lda-paper.tex

@@

\shrinkit
\subsection{ Future Work}
\shrinkit
\shrinkit

@@

% \section{Appendix}

>>>>

\documentclass[times, 10pt,twocolumn]{article} 
\usepackage{latex8}
\usepackage{times}
\usepackage{graphicx,amsmath,comment}
\newcommand{\igH}[1]{\includegraphics[height=.9\textheight]{#1}}
\newcommand{\igW}[1]{\includegraphics[width=.9\textwidth]{#1}}
\newcommand{\igWh}[1]{\includegraphics[width=.7\textwidth]{#1}}
\newcommand{\igWhalf}[1]{\includegraphics[width=.45\textwidth]{#1}}
\newcommand{\lda}{Latent Dirichlet Allocation}
\pagestyle{empty}

@@

\maketitle
\thispagestyle{empty}

\begin{abstract}

@@

\Section{Introduction}

We want to know about the topics of development that programmers
encountered while developing a software project. We want to know given
a certain period of time, what were the popular topics being worked on
during that period. We also want to know if any of those topics
occurred again, whether before or after our current period.  These
topics could help stakeholders determine what their coworkers were
working on, what were common issues, what requirements were being
worked on, etc.

Topic analysis extracts groups of words from documents (commit log
comments) that are word distributions, topics, that are combined to
characterize clusters of documents. These extracted topics often
correlate with actual development topics that the developers discussed
during the development of the software project.

Topic analysis often allows for a single document, such as a commit
message, to be related to multiple topics. This maps well to commits
to source code, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this case a topic
is a set of tokens extracted from commit messages found within a
projects source control system (SCS).

@@

Figure \ref{fig:lda} illustrates the kind of visualization we want to
automatically generate. We want to see topics that are unique to a
certain time period, displayed across the time axis. Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a single word drawn from its word
distribution.

In this paper we explore how topics shift over time in the source
control system (SCS) of a project, using several open source database
systems as examples. We analyze commit log comments in each time
window and we extract the topics of that time window. We expect that
topics will change over time and the similarities and differences in
these topics will indicate developer focus and changes in developer
focus as it changes over time. We will show that most topics are
locally relevant and thus a windowed topic analysis (local topic
analysis of windows of messages) can allow for both different kinds of
global and local analysis than global topic analysis would.

Our contributions include:
\begin{itemize}
\item We demonstrate that windowed topic analysis most topics are local
  in scope, thus would be missed by a global topic analysis.
\item We present multiple visualizations of topics and trends over time to aid
  communication and analysis of topics and trends.
\item We provide an exploratory case study using these techniques on
  several open source database systems.
\end{itemize}

@@

%%%   \Section{Introduction}

@@

\Section{Background}

@@

%\SubSection{Intro to LDA}

@@

\shrinkit
\SubSection{LDA, LSI and Semantic Clustering}
\shrinkit

@@

\shrinkit
\Section{Preliminary Case Study}
\shrinkit

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
examined the top 10 most frequent words in that topic.  We found there
were common words across topic clusters, such as \emph{diffs},
\emph{annotate} and \emph{history}; since these words occur so often
in the MySQL 3.23 messages, we probably should be viewed as stop
words. There were notable transitional topics such as in the first
window the word \emph{BitKeeper} appears because MySQL adopted
Bitkeeper for their source control system yet in the following windows
there were no mentions of BitKeeper. \emph{RENAME} also only appeared
once in the first window and never again. Per each topic we tried to
identify the purpose of the topic; to our surprise we found that even
with only a little familiarity with the code base that naming the
topic was straightforward.

@@

\shrinkit
\Section{Methodology}
\shrinkit

@@

\shrinkit
\SubSection{Extraction of  Repositories and Documents}
\shrinkit

@@

\shrinkit
\SubSection{Windowed Sets}
\shrinkit

@@

\shrinkit
\SubSection{Apply Topic Analysis to each Window}
\shrinkit

@@

%\SubSection{Extract Clusters}

%   6. [ ] Cluster Similarity [0/1]
\shrinkit
\SubSection{Topic Similarity}
\shrinkit
%      1. [ ] need the tool
%             grab from the cluster analysis
%   7. [ ] Continuous Cluster Analysis

@@

\shrinkit
\SubSection{Visualization}
\shrinkit
%       1. [ ] need the visualizer

@@

\shrinkit
\Section{Results}
\shrinkit

@@

%\SubSection{MySQL}
%XXX DO MYSQL?
\shrinkit
\SubSection{PostgreSQL}
\shrinkit
%   3.   mysql

@@

\shrinkit
\SubSection{Firebird}
\shrinkit
%   5.   what's that other db?

@@

\shrinkit
\SubSection{MaxDB 7.500}
\shrinkit

@@

The query optimizer was a topic that recurred during MaxDB's
development. In our plots, topics that mention \emph{optimizer} occur
four times, yet in the global-trend-view (Figure
\ref{fig:statictopics}) it is not in any of the topics. A query
optimizer is an important component of an DBMS, but as we have shown
it does not appear as a topic on its own. We tried to remove words to
see if we could get an \emph{optimizer} topic. After removing stop
words and then two of the most common words, the global analysis
finally had a topic with optimizer in its topic ten words. Our
analysis shows that optimizer was important but it had been obscured
by the global topic analysis, which used the entire history of
messages, but would have been noticed using the more local topic
analysis such as our windowed topic analysis, which used a window of
messages.

@@

\shrinkit
\SubSection{Compare with topics over the entire development}
\shrinkit

@@

%\SubSection{Validation}

%   1. well i'm stumped
%   2. look at it?
%   3. investigate those revisions and check the proportion etc
%   4. exploratory
%   5.   work it out
%8.   Validity Threats [0/5]
\shrinkit
\Section{Validity Threats}
\shrinkit

@@

\shrinkit
\Section{Conclusions}
\shrinkit
%11.   Start file file:/home/abez/projects/lda-paper/lda-paper.tex

@@

\shrinkit
\SubSection{ Future Work}
\shrinkit
\shrinkit

@@

% \Section{Appendix}
